Paper_Figure

### 2025-03-24 LMD-paper/paper_figure.Rmd
revision

---
title: "Paper_Figure"
author: "Ruiqi Li"
output: pdf_document
date: '2023-10-16'
---
# Load Data & Preprocess
```{r}
library("Seurat")
library("RColorBrewer")
library("dplyr")
library("patchwork")
library("ggplot2")
# library("ggsignif")
library("LocalizedMarkerDetector")
source("paper_figure_function.R", echo = F)
source("paths.R")
figure_path = file.path(dir.path0,"figures")
dir.create(figure_path,recursive = TRUE)
setwd(figure_path)
```

# SupFig - Bone Marrow Cell Embedding
```{r}
dir.path <- dir.path0
data_source = "tabular_muris"
folder.path <- file.path(dir.path,data_source)
tissue_name = "marrow_facs"
tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
cell_label = tiss$cell_ontology_class
coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
p1 = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)

tissue_name = "marrow_droplet"
tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
cell_label = tiss$cell_ontology_class
coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
p2 = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)

p = ((p1 / p2) + plot_annotation(tag_levels = 'A')) & theme(plot.tag = element_text(size = 20))
ggsave(filename = file.path(figure_path,"supfig-bonemarrow.png"), plot = p, width = 8, height = 10, dpi = 300)
```

@@

# Fig - the AUROC given the rank and the ground truth
```{r}
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_no_filter","hvg")
method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4","HVGs"),method_ls)
```
```{r}
dir.path <- dir.path0
data_source = "tabular_muris"
tissue_name = "marrow_facs"
folder.path = file.path(dir.path,"benchmark")
```
```{r}
# Define Full name
tissue_ls = list("tabular_muris" = c("marrow_facs","pancreas_facs","lung_facs"),
                 "azimuth" = c("human_kidney", "human_bone_marrow", "human_pancreas", "mouse_motor_cortex", "human_lung_hlca", "human_motor_cortex"))
data_source_full_name = c("tabular_muris" = "Tabular Muris","azimuth" = "Azimuth")
tissue_full_name = lapply(unlist(tissue_ls),function(i){
  tmp_name = 
  if(grepl("marrow",i)){tmp_name = "Bone Marrow"}
  else if(grepl("lung_hlca",i)){tmp_name = "Lung v2 (HLCA)"}
  else if(grepl("lung",i)){tmp_name = "Lung"}
  else if(grepl("pancreas",i)){tmp_name = "Pancreas"}
  else if(grepl("kidney",i)){tmp_name = "Kidney"}
  else if(grepl("motor_cortex",i)){tmp_name = "Motor Cortex"}
  else{tmp_name = NA}
  tmp_name
})
names(tissue_full_name) = unlist(tissue_ls)
tissue_full_name = unlist(lapply(1:length(tissue_full_name),function(i){
  ori_name = names(tissue_full_name)[i]
  if(grepl("_facs|mouse",ori_name)){
    tmp_name = paste0("Mouse-\n",tissue_full_name[i])
  }
  else if(grepl("human",ori_name)){
    tmp_name = paste0("Human-\n",tissue_full_name[i])
  }
  else{
    tmp_name = tissue_full_name[i]
  }
  tmp_name
}))
names(tissue_full_name) = unlist(tissue_ls)
```
```{r}
df_auc = do.call(rbind,
                lapply(1:length(tissue_ls),function(i){
                  df_all = NULL
                  for(j in tissue_ls[[i]]){
                    df = read.table(file.path(folder.path, names(tissue_ls)[i],paste0(j,"_auroc.csv")),header = TRUE)
                    df$'Tissue' = j
                    df_all = rbind(df_all,df)
                  }
                  df_all$'data_source' = names(tissue_ls)[i]
                  df_all}))
df_auc$Method = method_fullname[df_auc$Method]
df_auc = df_auc %>% filter(gt_set %in% c("CellMarkerDB","Top100"))
# Order Method based on mean AUROC
df_auc$Method <- with(df_auc, reorder(Method, -AUROC, FUN = median))
df_auc$gt_set <- factor(df_auc$gt_set,levels = c("CellMarkerDB","Top100"))
levels(df_auc$gt_set) = c("Criterion1-\n CellMarkerDB","Criterion2-\n MaxFC")
df_auc$Tissue = tissue_full_name[df_auc$Tissue]
df_auc$data_source = data_source_full_name[df_auc$data_source]
df_auc$data_source = factor(df_auc$data_source, levels = data_source_full_name)
```
## Figure
```{r}
coldef = setNames(
  colorRampPalette(brewer.pal(8, "Set1"))(length(method_fullname)),
  method_fullname)

@@

ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = fig, width = 12, height = 6)
```
## Table
```{r}
library("reshape2")
wide_data = df_auc; 
# wide_data$AUROC = round(wide_data$AUROC,3)
wide_data <- reshape2::dcast(wide_data, gt_set + Tissue + data_source ~ Method, value.var = "AUROC")
colnames(wide_data)[1:3] = c("Criterion","Tissue","DataSource")
wide_data[,1:3] = wide_data[,3:1]; colnames(wide_data)[1:3]=colnames(wide_data)[3:1]
wide_data <- wide_data[order(wide_data$DataSource,wide_data$Tissue,wide_data$Criterion),]
wide_data[,1:3] = apply(wide_data[,1:3],2,function(x) gsub("[\n]|Criterion1-|Criterion2-","",x))
wide_data[,1:2] = apply(wide_data[,1:2],2,function(x) {x[duplicated(x)] = "";x})
wide_data$Tissue[which(wide_data$Tissue == "Mouse-Motor Cortex")] = "Motor Cortex(Mouse)"
wide_data$Tissue[which(wide_data$Tissue == "Human-Motor Cortex")] = "Motor Cortex(Human)"
wide_data$Tissue = gsub("Human-|Mouse-","",wide_data$Tissue)

# calculate median rank
ranked_data <- data.frame(t(apply(wide_data[,4:ncol(wide_data)],1,function(x) rank(-x, na.last = "keep"))), check.names = FALSE)
ranked_data$Criterion = wide_data$Criterion
median_rank = ranked_data %>% group_by(Criterion) %>% summarize(across(1:(ncol(ranked_data)-1), ~ median(.x, na.rm = TRUE)))
median_rank$DataSource = c("Median Rank",""); median_rank$Tissue = "-"
wide_data = rbind(wide_data,median_rank)

@@

# write.csv(wide_data, file = file.path(folder.path,"AUROC_table.csv"), row.names = FALSE)
```
## Colored Table figure
```{r}
library(tidyr)

# separate wide_data into two by true_set
df_head = wide_data %>% filter(Criterion == " CellMarkerDB") %>% select(c("DataSource","Tissue"))
df_head[nrow(df_head),"DataSource"] = ""
df_head[nrow(df_head),"Tissue"] = "Median Rank"
df_head$DataSource = c(
  rep(df_head$DataSource[1],3),
  rep(df_head$DataSource[4],6),
  df_head$DataSource[10])

@@

df_long$Method = factor(df_long$Method, levels = rev(colnames(wide_data)[4:10]))
df_long$DataSource = factor(df_long$DataSource, levels = unique(df_long$DataSource))

# Plot the heatmap
p = ggplot(df_long, aes(x = Tissue, y = Method, fill = Rank)) +
    geom_tile(color = "white") + 
    scale_fill_distiller(palette = "YlOrBr", direction = -1, na.value = "white") +
    # scale_fill_gradient(low = "brown", high = "white", na.value = "white") +
    geom_text(aes(label = AUROC), size = 3) + 
    theme_minimal() + facet_grid(~ DataSource, scales = "free",space = "free", switch = "x") + ggtitle(paste0("True markerset ",i)) + 
    theme(plot.title = element_text(size = 15,face = "bold"),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
          strip.placement = 'outside',
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          panel.grid = element_blank(), 
          strip.text = element_text(size = 10,face = "bold"),
          axis.text.y = element_text(face = "bold", size = 10))

@@

p = wrap_plots(pl, nrow = 1) + plot_layout(guides = "collect")

ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = p, width = 12, height = 6)
```

@@

ggsave(filename = file.path(figure_path,"supfig-runtime.png"), plot = fig, width = 8, height = 5)
```
## Table
```{r}
df_runtime = df_runtime %>% filter(Method == "LMD") %>% arrange(Time, nCells, nGenes) %>% select(c("Time","nCells","nGenes")) %>% mutate(Time = round(Time, 2))
colnames(df_runtime)[1] = "Time (min)"
write.csv(df_runtime, file = file.path(folder.path,"RunTime_table.csv"), row.names = FALSE)
```

# Fig - Gene Ranking Comparisons between LMD and Other Methods
```{r}
dir.path <- dir.path0
data_source = "tabular_muris"
tissue_name = "marrow_facs"
df_benchmark = read.table(file.path(dir.path,"benchmark", data_source, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
rownames(df_benchmark) = df_benchmark$gene
tiss = readRDS(file.path(dir.path,data_source, paste0(tissue_name,".rds")))
method_vs = "lmd"
method_vs_set = c("wilcox_no_filter","hotspot","marcopolo","semi","haystack")
```
```{r}
sub = do.call(c,lapply(method_vs_set,function(method){
  sub = c(which(df_benchmark[,method] <= 100 & df_benchmark[,method_vs] > 500),
          which(df_benchmark[,method] > 500 & df_benchmark[,method_vs] <= 100) )
  sub
}))
df_label_all = df_benchmark[unique(sub),c(method_vs,method_vs_set)]
# df_label_all$'legend' = sapply(seq(nrow(df_label_all)),function(i){
#   paste(paste(method_fullname[colnames(df_label_all)],df_label_all[i,],sep = ":"),collapse = "\n")
# })
g = c("Cd19","Naaa","Mogat2","Grina","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb") # <100&>500
names(g) = rep(method_vs_set,each = 2)
pl = lapply(1:length(g),function(i){
  x = g[i]
  subtitle = sprintf("%s:%d, %s:%d",method_fullname[method_vs],df_label_all[x,method_vs],method_fullname[names(g)[i]],df_label_all[x,names(g)[i]])
  color_label = ifelse(df_label_all[x,method_vs] > df_label_all[x,names(g)[i]],"blue","red")
  FeaturePlot(tiss,features = x,order =TRUE) + labs(color = "Expression") + labs(title = x,subtitle = subtitle) + 
    theme(
      plot.title = element_text(face="bold", size = 30,color = color_label),
      plot.subtitle = element_text(size = 20),
      legend.title = element_text(size = 20),
      legend.text = element_text(size = 15),
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks=element_blank(),
      axis.line = element_blank())
  # + annotation_custom(grob = grid::textGrob(df_label_all[x,"legend"], x = 1, y = 1, just = c("right", "top"),gp = grid::gpar(col = "red", fontsize = 10)))
})
pl = pl[order(seq(length(g))%%2,decreasing = TRUE)]
pl[[1]] = pl[[1]] + labs(y = "Favored by LMD") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "red"))
pl[[(length(pl)/2)+1]] = pl[[(length(pl)/2)+1]] + labs(y = "Favored by Other methods") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "blue"))

fig = wrap_plots(pl,nrow = 2)

ggsave(filename = file.path(figure_path,"fig-ranking_comparison2.png"), plot = fig, width = 25, height = 10)
```
```{r}
pl_rank = lapply(method_vs_set,function(method){
  df_label = data.frame(df_benchmark[,c(method,method_vs)],label = rownames(df_benchmark))
  fullmethod = method_fullname[method]
  fullmethodvs = method_fullname[method_vs]
  df_label[c(which(df_label[,method] <= 100 & df_label[,method_vs] > 500)),"color"] = fullmethod
  df_label[c(which(df_label[,method] > 500 & df_label[,method_vs] <= 100) ),"color"] = fullmethodvs
  df_label$'color' = factor(df_label$'color',levels = c(fullmethodvs,fullmethod))
  rank_cor = stats::cor(log(df_label[,1:2],base = 10),method = "pearson")[1,2]
  colnames(df_label)[1:2] = c("x","y")
  p = ggplot(df_label, aes(x=x, y=y,color = color)) + geom_point(size = 1) + 
  scale_color_manual(values = c("red","blue","grey50"),
                     breaks = c(fullmethodvs,fullmethod),
                     labels = c("Favored by LMD","Favored by Other methods")) + 
    geom_abline(intercept = 0, slope = 1, col = "grey", linetype = "dashed") + 
    scale_x_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) +
    scale_y_log10(breaks = 10^(0:4), labels = expression(10^0, 10^1, 10^2, 10^3, 10^4)) + 
    expand_limits(x = c(1, 10^4), y = c(1, 10^4)) +
    labs(x = paste0(method_fullname[method]," Rank"), 
         y = paste0(method_fullname[method_vs]," Rank"),
         color = "Highlighted Genes",
         title = paste("cor =", round(rank_cor, 2)))
  # annotate gene labels
  p = p + ggrepel::geom_text_repel(data = df_label %>% filter(!is.na(color)) %>% filter(label %in% g), aes(label=label),min.segment.length = unit(0, 'lines'),nudge_y = -0.5,show.legend = FALSE,size = 5)
  p = p + theme(
    plot.title = element_text(size = 15),
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 15),
    axis.title.x = element_text(size = 20),
    axis.title.y = element_text(size = 20),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    panel.grid = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"))
})
for(i in 2:length(pl_rank)){
  pl_rank[[i]] <- pl_rank[[i]] + theme(axis.title.y=element_blank())
}
p = (wrap_plots(pl_rank,nrow = 1) & theme(legend.position="bottom")) + plot_layout(guides = "collect")

ggsave(filename = file.path(figure_path,"fig-ranking_comparison1.png"), plot = p, width = 25, height = 6)
```

>>>>

---
title: "Paper_Figure"
author: "Ruiqi Li"
output: pdf_document
date: '2023-10-16'
---
# Load Data & Preprocess
```{r}
library("Seurat")
library("RColorBrewer")
library("dplyr")
library("patchwork")
library("ggplot2")
# library("ggsignif")
# remotes::install_github("KlugerLab/LocalizedMarkerDetector")
# remotes::install_github("KlugerLab/GeneTrajectory")
library("LocalizedMarkerDetector")
source("paper_figure_function.R", echo = F)
source("paths.R")
figure_path = file.path(dir.path0,"figures")
if(!dir.exists(figure_path)){dir.create(figure_path,recursive = TRUE)}
```

# SupFig - Bone Marrow Cell Embedding
```{r}
dir.path <- dir.path0
data_source = "tabular_muris"
folder.path <- file.path(dir.path,data_source)
tissue_name = "marrow_facs"
tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
unique_labels <- unique(tiss$cell_ontology_class)
label_numbers <- setNames(seq_along(unique_labels), unique_labels)
tiss$cell_number <- as.character(label_numbers[tiss$cell_ontology_class])
Idents(tiss) <- tiss$cell_number 
coldef <- setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(label_numbers)), 
                   label_numbers)
p1 <- DimPlot(tiss, group.by = "cell_number", label = TRUE, cols = coldef) +
  guides(color = guide_legend(ncol = 1, override.aes = list(size = 3))) +
  scale_color_manual(values = coldef, labels = paste0(seq_along(unique_labels), " - ", unique_labels)) +  # Modify legend labels
  labs(title = NULL)

tissue_name = "marrow_droplet"
tiss <- readRDS(file.path(folder.path,paste0(tissue_name,".rds")))
unique_labels <- unique(tiss$cell_ontology_class)
label_numbers <- setNames(seq_along(unique_labels), unique_labels)
tiss$cell_number <- as.character(label_numbers[tiss$cell_ontology_class])
Idents(tiss) <- tiss$cell_number 
coldef <- setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(label_numbers)), 
                   label_numbers)
p2 <- DimPlot(tiss, group.by = "cell_number", label = TRUE, cols = coldef) +
  guides(color = guide_legend(ncol = 1, override.aes = list(size = 3))) +
  scale_color_manual(values = coldef, labels = paste0(seq_along(unique_labels), " - ", unique_labels)) +  # Modify legend labels
  labs(title = NULL)

p = ((p1 / p2) + plot_annotation(tag_levels = 'A')) & theme(plot.tag = element_text(size = 20))
ggsave(filename = file.path(figure_path,"supfig-bonemarrow.png"), plot = p, width = 9, height = 10, dpi = 300)
```

@@

# Fig - the AUROC given the rank and the ground truth
```{r}
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_no_filter","hvg", "cosg", "scmarker")
method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v4","HVGs", "COSG", "SCMarker"),method_ls)
```
```{r}
dir.path <- dir.path0
# data_source = "tabular_muris"
# tissue_name = "marrow_facs"
folder.path = file.path(dir.path,"benchmark")
```
```{r}
# Define Full name
tissue_ls = list("tabular_muris" = c("marrow_facs","pancreas_facs","lung_facs","marrow_droplet"),
                 "azimuth" = c("human_kidney", "human_bone_marrow", "human_pancreas", "mouse_motor_cortex", "human_lung_hlca", "human_motor_cortex"))
data_source_full_name = c("tabular_muris" = "Tabular Muris","azimuth" = "Azimuth")
tissue_full_name = lapply(unlist(tissue_ls),function(i){
  tmp_name = 
  if(grepl("marrow",i)){tmp_name = "Bone Marrow"}
  else if(grepl("lung_hlca",i)){tmp_name = "Lung v2 (HLCA)"}
  else if(grepl("lung",i)){tmp_name = "Lung"}
  else if(grepl("pancreas",i)){tmp_name = "Pancreas"}
  else if(grepl("kidney",i)){tmp_name = "Kidney"}
  else if(grepl("motor_cortex",i)){tmp_name = "Motor Cortex"}
  else{tmp_name = NA}
  tmp_name
})
names(tissue_full_name) = unlist(tissue_ls)
tissue_full_name = unlist(tissue_full_name)
tissue_full_name["human_motor_cortex"] = paste0(tissue_full_name["human_motor_cortex"],"(Human)")
tissue_full_name["mouse_motor_cortex"] = paste0(tissue_full_name["mouse_motor_cortex"],"(Mouse)")

# add technologies as suffix
techs = c(rep("FACS",3),"Droplet","10x","Cite-seq","Smart-seq2","10x","10x","Smart-seq v4")
tissue_full_name = setNames(paste0(tissue_full_name," ","(",techs,")"),names(tissue_full_name))
```
```{r}
df_auc = do.call(rbind,
                lapply(1:length(tissue_ls),function(i){
                  df_all = NULL
                  for(j in tissue_ls[[i]]){
                    df = read.table(file.path(folder.path, names(tissue_ls)[i],paste0(j,"_auroc.csv")),header = TRUE)
                    df$'Tissue' = j
                    df_all = rbind(df_all,df)
                  }
                  df_all$'data_source' = names(tissue_ls)[i]
                  df_all}))
```
## Louvain resolution for Cluster-dependent methods
```{r}
# Plot performance for each datasets
method_subls = c("wilcox_no_filter","cosg")
df_auc_sub = do.call(rbind,lapply(method_subls, function(method){
  df = do.call(rbind,
        lapply(1:length(tissue_ls),function(i){
            df_all = NULL
            for(j in tissue_ls[[i]]){
                df = read.table(file.path(folder.path, names(tissue_ls)[i],method,paste0(j,"_auroc.csv")),header = TRUE)
                df$'Tissue' = j
                df_all = rbind(df_all,df)
            }
            df_all$'data_source' = names(tissue_ls)[i]
            df_all}))
  df$'Method' = method
  df
}))
```
```{r}
# plot louvain-clustering vs. real celltype
df_tmp = do.call(rbind,
        lapply(1:length(tissue_ls),function(i){
            df_all = NULL
            for(j in tissue_ls[[i]]){
                df = read.table(file.path(folder.path, names(tissue_ls)[i],paste0(j,"_louvain_clusters.csv")),header = TRUE)
                df$'Tissue' = j
                df_all = rbind(df_all,df)
            }
            df_all$'data_source' = names(tissue_ls)[i]
            df_all}))
df_tmp$Tissue = tissue_full_name[df_tmp$Tissue]
df_tmp <- df_tmp %>%
  mutate(Tissue = factor(Tissue, levels = unique(Tissue[rev(order(data_source))])))

p = df_tmp %>% ggplot(aes(x = as.factor(res), y = ncluster, group = 1)) + 
  geom_line() + geom_hline(aes(yintercept = ncelltype), color="red", linetype="dashed") + 
    facet_wrap(~Tissue, scales = "free") + xlab("Resolution") + ggtitle("The number of Louvain clusters by different resolution")
ggsave(filename = file.path(figure_path,"supfig-seurat_performance1.png"), 
       plot = p, width = 10, height = 5)
```
```{r}
# plot Seurat
df_tmp = df_auc_sub %>% filter(gt_set %in% c("CellMarkerDB","Top100"))
df_tmp$gt_set = as.factor(df_tmp$gt_set); levels(df_tmp$gt_set) = c("Trueset1","Trueset2")
df_tmp$Tissue = tissue_full_name[df_tmp$Tissue]
df_tmp <- df_tmp %>%
  mutate(Tissue = factor(Tissue, levels = unique(Tissue[rev(order(data_source))])))

pl = lapply(method_subls,function(method){
  df_tmp %>% filter(Method == method) %>% ggplot(aes(x = as.factor(res), y = AUROC, group = gt_set, color = gt_set)) + 
  geom_line() + 
    facet_wrap(~Tissue) +
    ggtitle(method_fullname[method]) + xlab("Resolution")
})
p = wrap_plots(pl, nrow = 2)
ggsave(filename = file.path(figure_path,"supfig-seurat_performance.png"), 
       plot = p, width = 16, height = 16)
```
```{r}
# # Replace cluster-dependent method w/ resolution of maximum performance
# df_auc_sub %>% 
#   filter(gt_set %in% c("Top100", "CellMarkerDB")) %>% 
#   group_by(res) %>% 
#   summarize(median_AUROC = median(AUROC, na.rm = TRUE)) %>% 
#   arrange(desc(median_AUROC))
# 
# default_res = 5
# # df_auc_sub = df_auc_sub %>%
# #     group_by(Tissue, gt_set, data_source, Method) %>%
# #     filter(AUROC == max(AUROC))
# df_auc_sub = df_auc_sub %>% filter(res == 5) %>% select(colnames(df_auc))
# '%notin%' <- Negate('%in%')
# df_auc = rbind(df_auc %>% filter(Method %notin% method_subls),df_auc_sub)
```
## Re-organize table
```{r}
# re-organize the df
# Order Method based on mean AUROC
df_auc$Method = method_fullname[df_auc$Method]
top_name = "Top100"
df_auc = df_auc %>% filter(gt_set %in% c("CellMarkerDB",top_name))
df_auc$Method <- with(df_auc, reorder(Method, -AUROC, FUN = median))
df_auc$gt_set <- factor(df_auc$gt_set,levels = c("CellMarkerDB",top_name))
levels(df_auc$gt_set) = c("Criterion1-\n CellMarkerDB","Criterion2-\n MaxFC")
df_auc$Tissue = tissue_full_name[df_auc$Tissue]
df_auc$data_source = data_source_full_name[df_auc$data_source]
df_auc$data_source = factor(df_auc$data_source, levels = data_source_full_name)
```

## Figure
```{r}
coldef = setNames(
  colorRampPalette(brewer.pal(8, "Set1"))(length(method_fullname)),
  method_fullname)

@@

ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = fig, width = 12, height = 6)
```
## Table
```{r}
library("reshape2")
wide_data = df_auc; 
# wide_data$AUROC = round(wide_data$AUROC,3)
wide_data <- reshape2::dcast(wide_data, gt_set + Tissue + data_source ~ Method, value.var = "AUROC")
colnames(wide_data)[1:3] = c("Criterion","Tissue","DataSource")
wide_data[,1:3] = wide_data[,3:1]; colnames(wide_data)[1:3]=colnames(wide_data)[3:1]
wide_data <- wide_data[order(wide_data$DataSource,wide_data$Tissue,wide_data$Criterion),]
wide_data[,1:3] = apply(wide_data[,1:3],2,function(x) gsub("[\n]|Criterion1-|Criterion2-","",x))
wide_data[,1:2] = apply(wide_data[,1:2],2,function(x) {x[duplicated(x)] = "";x})

# calculate median rank
ranked_data <- data.frame(t(apply(wide_data[,4:ncol(wide_data)],1,function(x) rank(-x, na.last = "keep", ties.method = "min"))), check.names = FALSE)
ranked_data$Criterion = wide_data$Criterion
median_rank = ranked_data %>% group_by(Criterion) %>% summarize(across(1:(ncol(ranked_data)-1), ~ median(.x, na.rm = TRUE)))
median_rank$DataSource = c("Median Rank",""); median_rank$Tissue = "-"
wide_data = rbind(wide_data,median_rank)

@@

# write.csv(wide_data, file = file.path(folder.path,"AUROC_table.csv"), row.names = FALSE)
```
### Colored Table figure
```{r}
library(tidyr)

# separate wide_data into two by true_set
df_head = wide_data %>% filter(Criterion == " CellMarkerDB") %>% select(c("DataSource","Tissue"))
df_head[nrow(df_head),"DataSource"] = ""
df_head[nrow(df_head),"Tissue"] = "Median Rank"
df_head$DataSource = c(
  rep(df_head$DataSource[1],4),
  rep(df_head$DataSource[5],6),
  df_head$DataSource[nrow(df_head)])

@@

df_long$Method = factor(df_long$Method, levels = rev(colnames(wide_data)[-c(1:3)]))
df_long$DataSource = factor(df_long$DataSource, levels = unique(df_long$DataSource))
tissue_order = unique(df_long$Tissue[order(df_long$DataSource)]); tissue_order[1:4] = tissue_order[c(2:4,1)]
df_long <- df_long %>%
  mutate(Tissue = factor(Tissue, levels = tissue_order))

# Plot the heatmap
p = ggplot(df_long, aes(x = Tissue, y = Method, fill = Rank)) +
    geom_tile(color = "white") + 
    scale_fill_distiller(palette = "YlOrBr", direction = -1, na.value = "white", limits = c(1, max(df_long$Rank, na.rm = TRUE)), 
                         breaks = seq(1, max(df_long$Rank, na.rm = TRUE))) +
    # scale_fill_gradient(low = "brown", high = "white", na.value = "white") +
    geom_text(aes(label = AUROC), size = 3) + 
    theme_minimal() + facet_grid(~ DataSource, scales = "free",space = "free", switch = "x") + ggtitle(paste0("True markerset ",i)) + 
    theme(plot.title = element_text(size = 15,face = "bold"),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
          strip.placement = 'outside',
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          panel.grid = element_blank(), 
          strip.text = element_text(size = 10,face = "bold"),
          axis.text.y = element_text(face = "bold", size = 10))

@@

p = wrap_plots(pl, nrow = 1) + plot_layout(guides = "collect")
```
```{r}
ggsave(filename = file.path(figure_path,"fig-known_marker_comparison.png"), plot = p, width = 13, height = 8)
```

@@

ggsave(filename = file.path(figure_path,"supfig-runtime.png"), plot = fig, width = 8, height = 5)
```
## Table
```{r}
df_runtime = df_runtime %>% filter(Method == "LMD") %>% arrange(Time, nCells, nGenes) %>% select(c("data_source","Tissue","nCells","ncluster","nGenes","Time")) %>% mutate(Time = round(Time, 2))
colnames(df_runtime)[grep("Time",colnames(df_runtime))] = "Time (min)"
write.csv(df_runtime, file = file.path(folder.path,"RunTime_table.csv"), row.names = FALSE)
```

# Fig - Gene Ranking Comparisons between LMD and Other Methods
```{r}
method_vs = "lmd"
method_vs_set = c("wilcox_no_filter","cosg","hotspot","marcopolo","scmarker","semi","haystack")
```
## Selected genes Table
### Bone marrow facs
```{r}
dir.path <- dir.path0
data_source = "tabular_muris"
tissue_name = "marrow_facs"
df_benchmark = read.table(file.path(dir.path,"benchmark", data_source, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
rownames(df_benchmark) = df_benchmark$gene
tiss = readRDS(file.path(dir.path,data_source, paste0(tissue_name,".rds")))

```
```{r}
sub = do.call(c,lapply(method_vs_set,function(method){
  sub = c(which(df_benchmark[,method] <= 100 & df_benchmark[,method_vs] > 500),
          which(df_benchmark[,method] > 500 & df_benchmark[,method_vs] <= 100) )
  sub
}))
df_label_all = df_benchmark[unique(sub),c(method_vs,method_vs_set)]
# df_label_all$'legend' = sapply(seq(nrow(df_label_all)),function(i){
#   paste(paste(method_fullname[colnames(df_label_all)],df_label_all[i,],sep = ":"),collapse = "\n")
# })
# g = c("Cd19","Naaa","Mogat2","Grina","Abca13","Zf12","Il2rb","Phgdh","Vpreb2","Fyb") # <100&>500 # before revision

g = c("Hp","Cyp11a1",#wilcox
      "1810033B17Rik","Pltp",#cosg
      "Mogat2","Grina",#hotspot
      "Abca13","Zf12",#marcopolo
      "Pilra","Rgs14",#scmarker
      "Il2rb","Phgdh",#semi
      "Vpreb2","Fyb"#schaystack
      ) # <100&>500

df_label = df_label_all[g,]
df_label = df_label[order(df_label$lmd),]
colnames(df_label) = method_fullname[colnames(df_label)]
write.csv(df_label,file = file.path(dir.path,"benchmark", paste0(tissue_name,"_benchmark_rank_table_selectedgenes.csv")),row.names = TRUE)
```
### human kidney 10x
```{r}
dir.path <- dir.path0
data_source = "azimuth"
tissue_name = "human_kidney"
df_benchmark = read.table(file.path(dir.path,"benchmark", data_source, paste0(tissue_name,"_benchmark_rank_table.csv")),header = TRUE)
rownames(df_benchmark) = df_benchmark$gene
tiss = readRDS(file.path(dir.path,data_source, paste0(tissue_name,".rds")))
tiss = RunUMAP(tiss, dims = 1:20, seed.use = 42)
```
```{r}
# Plot DimPlot
cell_label = tiss$cell_ontology_class
coldef = setNames(colorRampPalette(brewer.pal(12, "Paired"))(length(unique(cell_label))),unique(cell_label))
p1 = DimPlot(tiss, group.by = "cell_ontology_class",label = FALSE, cols = coldef) + guides(color = guide_legend(ncol = 1,override.aes = list(size = 3))) + labs(title = NULL)
```
```{r}
sub = do.call(c,lapply(method_vs_set,function(method){
  sub = c(which(df_benchmark[,method] <= 100 & df_benchmark[,method_vs] > 500),
          which(df_benchmark[,method] > 500 & df_benchmark[,method_vs] <= 100) )
  sub
}))
df_label_all = df_benchmark[unique(sub),c(method_vs,method_vs_set)]
# kidney
g = c("SPINK8","SPP1",#wilcox
      "RAMP2","KIAA0101",#cosg
      "SAA2","GSTA1",#hotspot
      "MEIS2","AIM1",#scmarker
      "MUC16","ATF3",#semi
      "CEACAM6","COX7A1"#schaystack
      )

df_label = df_label_all[g,]
df_label = df_label[order(df_label$lmd),]
colnames(df_label) = method_fullname[colnames(df_label)]
write.csv(df_label,file = file.path(dir.path,"benchmark", paste0(tissue_name,"_benchmark_rank_table_selectedgenes.csv")),row.names = TRUE)
```
## FeaturePlot
```{r}
names(g) = rep(method_vs_set,each = 2)
unified_range <- range(FetchData(tiss, vars = g), na.rm = TRUE)

pl = lapply(1:length(g),function(i){
  x = g[i]
  subtitle = sprintf("%s:%d, %s:%d",method_fullname[method_vs],df_label_all[x,method_vs],method_fullname[names(g)[i]],df_label_all[x,names(g)[i]])
  color_label = ifelse(df_label_all[x,method_vs] > df_label_all[x,names(g)[i]],"blue","red")
  FeaturePlot(tiss,features = x,order =TRUE,pt.size = 1) + labs(color = "Expression") + scale_color_gradientn(colors = c("lightgrey", "blue"), 
                          limits = unified_range) + labs(title = x,subtitle = subtitle) + 
    theme(
      plot.title = element_text(face="bold", size = 30,color = color_label),
      plot.subtitle = element_text(size = 20),
      legend.title = element_text(size = 20),
      legend.text = element_text(size = 15),
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks=element_blank(),
      axis.line = element_blank())
  # + annotation_custom(grob = grid::textGrob(df_label_all[x,"legend"], x = 1, y = 1, just = c("right", "top"),gp = grid::gpar(col = "red", fontsize = 10)))
})
pl = pl[order(seq(length(g))%%2,decreasing = TRUE)]
pl[[1]] = pl[[1]] + labs(y = "Favored by LMD") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "red"))
pl[[(length(pl)/2)+1]] = pl[[(length(pl)/2)+1]] + labs(y = "Favored by Other methods") + theme(axis.title.y = element_text(size = 20,face="bold", colour = "blue"))

fig = wrap_plots(pl,nrow = 2) + plot_layout(guides = "collect")
